---
title: "Welcome to some of my data visualization projects"
format:
  html:
    embed-resources: true
    link-external-newwindow: true
    page-layout: full
    toc-depth: 2
engine: knitr
---


# Visulization of American Housing
## The Homelessness Crisis
::: {.columns} 
::: {.column width="35%"}
<div style="padding-right: .25rem;">
### Individual Emergency Shelter Beds Per Homeless Individual
When looking at emergency shelter capacity relative to the homeless population, clear regional disparities emerge. States shown in red face shortages, with fewer than one shelter bed per homeless individual, while blue states indicate relative higher capacity. This visualization highlights structural differences in how states address homelessness and emergency housing needs. The data used for this visualization was gathered from HUD data[^1] and a yearly report created by the World Population review.[^2] 
</div>
:::

::: {.column width="65%"}
<div style="padding-left: .25rem;">
```{r}
#| echo: false
#| message: false
#| warning: false

packages <- c("tidyverse", "ggplot2", "plotly", "usmap")
invisible(lapply(packages, library, character.only = TRUE))
rm(packages)

HomeState <- read.csv("VizCode/HomPopState.csv")

HomeStateUse <- HomeState |>
  mutate(total = rowSums(across(4:11), na.rm = TRUE),
         BedPer = total / HPop,
         region = tolower(State))

states_map <- map_data("state") |>
  left_join(HomeStateUse, by = "region")

p <- ggplotly(
  ggplot(states_map, aes(
    long, lat, group = group,
    fill = BedPer,
    text = paste0(
      "State: ", tools::toTitleCase(region),
      "<br>Emergency Shelter Beds per Homeless Individual: ",
      round(BedPer, 2)
    )
  )) +
    geom_polygon(color = "black", linewidth = 0.3) +
    coord_fixed(1.3) +
    scale_fill_gradient2(
      low = "#570C18",
      high = "#000087",
      midpoint = 1,
      labels = scales::number_format(accuracy = 0.01),
      name = "Beds per Individual"
    ) +
    labs(
      title = "Emergency Shelter Beds per Homeless Individual (2024)",
      subtitle = "Red = shortage (<1), Blue = surplus (>1)"
    ) +
    theme_void()+
  theme(
    panel.background = element_rect(fill = "lightblue", colour = "lightblue"),
    plot.background = element_rect(fill = "lightblue"),
    legend.position = "bottom"
  ),
  tooltip = "text"
) 
  
# Find the trace that has a marker color scale
trace_id <- which(
  sapply(p$x$data, function(x) !is.null(x$marker$colorbar))
)

# Modify the colorbar 
p$x$data[[trace_id]]$marker$colorbar <- list(
  orientation = "h",
  x = 0.5,
  xanchor = "center",
  y = -0.15,
  thickness = 15,
  title = list(
    text = "Beds per Individual",
    side = "bottom"
  )
)

p

 rm(HomeState,HomeStateUse,p,states_map, trace_id)

```
</div>
:::
:::

# Visulization of The American Healthcare Crisis


### Medicaid Adult Coverage Expansion and Debt

:::{.columns}
::: {.column width ="65%"}

```{r}
#| label: MedicadeExpansion
#| echo: false
#| message: false
#| warning: false
MDEdf <- read.csv("VizCode/MedicalDebtExpansion.csv")
MDEdf <- MDEdf |>
  mutate(region = state, .before = time_period) |> 
  select(!state & !time_period & !cohort &!population) |> 
  mutate(Expanded = as.character(Expanded),
  Expanded = case_when(
    Expanded == 1 ~ "Medicaid Expanded",
    Expanded == 0 ~ "Medicaid Not Expanded",
    TRUE ~ Expanded
  ))

MDEdesc <- MDEdf |> 
  group_by(Expanded) |>
  filter(!is.na(point_estimate)) |> 
  summarize(
    Count = n(),
    `Mean` = mean(point_estimate),
    `Standard Deviation` = sd(point_estimate),
    Range =str_c(min(point_estimate)," : ",(max(point_estimate)))
  ) |> 
  na.omit()

ggplot(MDEdesc, aes(x = Expanded, y = Mean, fill = Expanded)) +
  
  # Bars
  geom_col(width = 0.55, alpha = 0.9) +
  
  # National average reference line
  geom_hline(
    yintercept = 5.04,
    linetype = "dashed",
    linewidth = 0.5,
    color = "gray30"
  ) +
  
  # Annotation for national average (to the right)
  annotate(
    "text",
    x = Inf,
    y = 5.04,
    label = "National Average: 5.04%",
    hjust = 3.35,
    vjust = -0.6,
    size = 4,
    color = "gray30"
  ) +
  
  # Labels
  labs(
    title = "Medical Debt in Collections by Medicaid Expansion",
    subtitle = "Share of individuals with a credit bureau record (%)",
    x = NULL,
    y = "Percent with Medical Debt in Collections",
    caption = "Source: The Commonwealth Fund & Medicaid.gov"
  ) +
  
  # Color palette
  scale_fill_manual(
    values = c(
      "Medicaid Expanded" = "#2C7FB8",
      "Medicaid Not Expanded" = "#D95F02"
    )
  ) +
  
  # Percent formatting
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0, 0.1))
  ) +
  
  # THIS fixes clipping
  coord_cartesian(clip = "off") +
  
  # Theme tuning for Quarto/web
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 13, color = "gray40"),
    axis.text.x = element_text(face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.caption = element_text(size = 10, color = "gray50"),
    plot.margin = margin(10, 45, 10, 10)
  )

rm(MDEdesc)
```

:::

:::{.column width = "35%"}
In this next section I will be investigating the differences in the COVID-19 Medicaid subsidies expansion and medical debt. In this bar graph Medicaid expansion data was gathered from Medicaid.com [^3] and The Commonwealth Fund [^4]
:::

:::












[^1]: [HUD Continuum of Care Homeless Assistance Programs Housing Inventory Count Report (2024)](https://files.hudexchange.info/reports/published/CoC_HIC_NatlTerrDC_2024.pdf)
[^2]: [World Population Review - Homeless Population By State (2025)](https://worldpopulationreview.com/state-rankings/homeless-population-by-state)
[^3]: [Medicaid Expansion Program Data (2023)](https://www.medicaid.gov/medicaid/program-information/downloads/medicaid-expansion-state-map.pdf)
[^4]: [The Commonwealth Fund - People With Medical Debt in Collections (2023)](https://www.commonwealthfund.org/datacenter/people-medical-debt-collections)
